A Gentle Introduction to Structural Causal Models

### 2021-05-11 scm.Rmd
scm

---
title: "scm_presentation"
author: "Daniel Saggau"
date: "5/11/2021"
output: beamer_presentation
---

@@

# Introduction

@@

# Fundamental Differences to begin

SCM:

- is a nonparametric SEM 
- has functional form rather than using probabilities
- entails features from the PO framework and graphical representation 
- conflict whether to use graphs or not

* A SEM is a parametric specification used in applied sciences (parameters contested)
* A Bayesian causal network is another popular causal model using conditional probabilities and NO functions 
* Differences in performance between BCN and SCM

#

@@

@@

>>>>

---
title: "A Gentle Introduction to Structural Causal Models"
author: "Daniel Saggau"
date: "5/11/2021"
output: beamer_presentation
---

@@

# Introduction (1)

- Machine learning has provided many insights into different research problems 
- One issue is the consideration of 'What are we actually predicting'?
- Mainstream tools are build on association-based learning 
- Associations are not enough for high stake settings 
- In disciplines like psychology or economics people are less interested in associational learning
- We want causation and not correlation

# Introduction (2)

@@

# History

* Path Analysis -> SEM -> SCM
*

# How to build a SCM

- Consists of system of equations 
- Assignment equation ':=' rather than regular equation '=' 
- is a nonparametric SEM 
- has functional form rather than using probabilities
- entails features from the PO framework and graphical representation 
- Exogenous factors are part of the model specification

# SCM Applications:

- Flexible simulations for higher order problems (intervention, counterfactual)
- Graphical visualization via directed acyclic graph
-

@@

# Fundamental Differences (1)

- conflict whether to use graphs or not 
* A SEM is a parametric specification used in applied sciences (parameters contested)
* A Bayesian causal network is another popular causal model using conditional probabilities and NO functions 
* Differences in performance between BCN and SCM

# Performance Evaluation

# Assumptions

**Error terms**

Regression: Omittable outside factor 
SCM/SEM; Latent influencial factor that is pivotal for the model specification but not observable

@@

# Causality -In Between Physical and Social Sciences

# Causality and Time

- Time in Physical Sciences: Mechanical and exact 
- Time in Social Sciences: Often Vague 
- Regular Time Specification is also more vague 
- To accomodate that issue, research on differential equation based SCMs started
